---
title: "ExamenRic_A01366889"
author: "Sebastian"
date: "04/04/2023"
output: html_document
---

## Primera parte SQL

# Ejercicio 1

SELECT Address
FROM Customers;


# Ejercicio 2


SELECT DISTINCT City
FROM Customers;


# Ejercicio 3


SELECT * FROM Customers
WHERE City = 'Berlin' OR City = 'Madrid';


# Ejercicio 4


SELECT * FROM Customers
LEFT JOIN Orders
ON Customers.CustomerID = Orders.CustomerID;


# Ejercicio 5

SELECT COUNT(CustomerID) AS NumClientes, City FROM Customers
GROUP BY City
ORDER BY COUNT(CustomerID) DESC;


## Segunda parte Data Viz


```{r}

# Ejercicio 1

library(dplyr)
library(readr)
library(tidyr)
library(stringr)
library(ggplot2)

library (tidyverse)
read_csv("C:/Users/Tanca/Documents/R/Data_Transformation_Viz/data/Market_size.csv")


# Ejercicio 2

library(data.table)
library(dplyr)

market <- fread("C:/Users/Tanca/Documents/R/Data_Transformation_Viz/data/Market_size.csv", skip = 5, sep = ",", header = TRUE)

market <- as.data.table (type.convert(market, as.is = TRUE))

precios_mx <- market %>%
  select (Geography, Category, `2023`) %>%
  filter (Geography == "Mexico")

precios_mx_gb <- precios_mx %>%
  group_by (Category) %>%
  summarise (mean_price_2023 = mean(`2023`, na.rm = TRUE))

precios_mx_gb


#Ejercicio 3

pl <- market %>%
  mutate (across(`2017`: `2023`, as.numeric)) %>%
  group_by(Geography, Category)


#Ejercicio 4

pl_viz <- market %>%
  mutate (across(`2017`: `2023`, as.numeric)) %>%
  filter (`Current Constant` == "Current Prices") %>%
  filter (str_sub (Geography, 1, 1) == "B") %>%
  group_by (Category) 

pl_viz <- gather(pl_viz, key = "Año", value = "Precio", (6:12))

ggplot (data = pl_viz, aes(x = Año, y = Precio, color = Geography, fill = Geography)) + 
  geom_bar (stat = "identity", position = position_stack(reverse = TRUE)) +
  scale_color_brewer (palette = "Dark2") +
  scale_fill_brewer (palette = "Dark2") +
  theme(axis.text.x = element_text(angle = 50, vjust = 0.5, hjust = 1)) +
  labs (title = "Precio promedio por año y país",
       x = "Año", y = "Precio por litro")


#Ejercicio 5

ggplot (pl_viz, aes(x = Año, y = Precio)) +
  geom_tile() +
  scale_fill_gradient(low = "orange", high = "red") +
  labs(title = "Heatmap de Precio vs Año", x = "Año", y = "Precio") +
  theme_classic()





```





























```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
